---
title: "Week 2 exercises"
author: "Wing-Ho Ko"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading common packages

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(janitor)
library(ggplot2)
```

## Exercise 1: plotting the penguins dataset

1. Make a plot of the Palmer penguin dataset, with flipper length on the horizontal axis and body mass on the vertical axis. Use color and shape to distinguish between different **species** of penguins

2. Focus only on the Adelie penguins. Plot flipper length on the horizontal axis and bill length on the vertical axis, and use size to indicate body mass. Use also color and shape to distinguish between penguins from different **islands**

3. Make a box plot of the body mass of the penguins as a function of species

## Exercise 2: biomass power-law scaling

Hatton et. al. (doi:10.1126/science.aac6284) found that predator biomass and prey biomass follows a power-law relationship across broad communities. Part of the data they used, for the case of Atlantic Ocean phytoplankton and microzooplankton, are extracted to the file marine_prey_predator_biomass.csv under the data folder.

Using the said data, make a scatter plot of predator (y) versus prey (x) biomass on a log-log scale. Include the x and y axes labels, and make sure that the unit has proper superscript. Also provide a suitable title to the plot.

## Exercise 3: NO₃⁻ and chlorophyll concentrations

The data file `mercator_no3_chl.csv` contains NO₃⁻ concentration (in mmol m⁻³) and chlorophyll concentration (in mg m⁻³), measured in June 2025, averaged over longitude between 150°W and 130°W, as functions of latitude and depth.

1. Make a raster plot (using `geom_rect()`, since the depths are uneven) of chlorophyll concentration as function of latitude (x) and depth (y). Use the usual oceanographic convention that depth increases downward, and use a sensible color scheme.

Hint: in your `gemo_rect()` call you will need to map the aesthetics `xmin`, `xmax`, `ymin`, and `ymax`. The corresponding variables have been calculated and included in the csv file.

2. Make a second raster plot with the same settings except making the color scale logarithmic

## Exercise 4: plotting the penguins dataset redux

Going back to the penguins dataset, make a figure consisting of 9 panels: 3 columns that indicate different species and 3 rows that indicate different islands. On each panel plots bill length against body mass. Make sure that all panels share the same set of x and y axes

## (Optional) Feedback questions

1. Are there any topic(s) from this week that still confuse you?

2. Which activities work and which doesn't? What do you want to see change and what do you want to keep?

